---
title: "A Market Segmentation and Purchase Drivers Process"
author: "T. Evgeniou"
output:
  html_document:
    css: ../../AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: ../../AnalyticsStyles/default.sty
always_allow_html: yes
---

<!-- **Note:** Assuming the working directory is "MYDIRECTORY/INSEADAnalytics" (where you have cloned the course material), you can create an html file by running in your console the command (pay attention to the directory you are in, check it using getwd()) rmarkdown::render("CourseSessions/InClassProcess/MarketSegmentationProcessInClass.Rmd") -->

> **IMPORTANT**: Please make sure you create a copy of this file with a customized name, so that your work (e.g. answers to the questions) is not over-written when you pull the latest content from the course github. 
This is a **template process for market segmentation based on survey data**, using the  [Boats cases A](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-A-prerelease.pdf) and  [B](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-B-prerelease.pdf).

All material and code is available at the INSEAD Data Science for Business website and GitHub. Before starting, make sure you have pulled the [course files](https://github.com/InseadDataAnalytics/INSEADAnalytics) on your GitHub repository. As always, you can use the `help` command in Rstudio to find out about any R function (e.g. type `help(list.files)` to learn what the R function `list.files` does).

**Note:** you can create an html file by running in your console the command 
rmarkdown::render("CourseSessions/InClassProcess/MarketSegmentationProcessInClass.Rmd") 
(see also a [potential issue with plots](https://github.com/InseadDataAnalytics/INSEADAnalytics/issues/75))

<hr>\clearpage

# The Business Questions

This process can be used as a (starting) template for projects like the one described in the [Boats cases A](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-A-prerelease.pdf) and  [B](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-B-prerelease.pdf). For example (but not only), in this case some of the business questions were: 

- What are the main purchase drivers of the customers (and prospects) of this company? 

- Are there different market segments? Which ones? Do the purchase drivers differ across segments? 

- What (possibly market segment specific) product development or brand positioning strategy should the company follow in order to increase its sales? 

See for example some of the analysis of this case in  these slides: <a href="http://inseaddataanalytics.github.io/INSEADAnalytics/Sessions2_3 Handouts.pdf"  target="_blank"> part 1</a> and <a href="http://inseaddataanalytics.github.io/INSEADAnalytics/Sessions4_5 Handouts.pdf"  target="_blank"> part 2</a>.

<hr>\clearpage

# The Process

The "high level" process template is split in 3 parts, corresponding to the course sessions 7-8, 9-10, and an optional last part: 

1. *Part 1*: We use some of the survey questions (e.g. in this case the first 29 "attitude" questions) to find **key customer descriptors** ("factors") using *dimensionality reduction* techniques described in the [Dimensionality Reduction](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html) reading of Sessions 7-8.

2. *Part 2*: We use the selected customer descriptors to **segment the market** using *cluster analysis* techniques described in the [Cluster Analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions45/ClusterAnalysisReading.html) reading of Sessions 9-10.

3. *Part 3*: For the market segments we create, we will use *classification analysis* to classify people based on whether or not they have purchased a product and find what are the **key purchase drivers per segment**. For this part we will use [classification analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/ClassificationProcessCreditCardDefault.html) techniques.

Finally, we will use the results of this analysis to make business decisions, e.g. about brand positioning, product development, etc., depending on our market segments and key purchase drivers we find at the end of this process.

```{r setuplibraries, echo=FALSE, message=FALSE}
suppressWarnings(source("../../AnalyticsLibraries/library.R"))
# Package options
#suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')

# Run below only once, then comment out
# New versions of the networkD3 package may not work properly, so install the following version
#packageurl <- "https://cran.r-project.org/src/contrib/Archive/networkD3/networkD3_0.2.13.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")
```

<hr>\clearpage

# The Data

First we load the data to use (see the raw .Rmd file to change the data file as needed):

```{r setupdata1E, echo=TRUE, tidy=TRUE}
# Please ENTER the name of the file with the data used. The file should be a .csv with one row per observation (e.g. person) and one column per attribute. Do not add .csv at the end, make sure the data are numeric.
datafile_name = "../Sessions23/data/Boats.csv"

# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE = 0.5

# Please enter the maximum number of observations to show in the report and slides. 
# DEFAULT is 10. If the number is large the report may be slow.
max_data_report = 10
```

```{r}
ProjectData <- read.csv(datafile_name)
ProjectData <- data.matrix(ProjectData) 
ProjectData_INITIAL <- ProjectData

```

<hr>\clearpage

# Part 1: Key Customer Characteristics

The code used here is along the lines of the code in the reading  [FactorAnalysisReading.Rmd](https://github.com/InseadDataAnalytics/INSEADAnalytics/blob/master/CourseSessions/Sessions23/FactorAnalysisReading.Rmd). We follow the process described in the [Dimensionality Reduction ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html) reading. 

In this part we also become familiar with:

1. Some visualization tools;
2. Principal Component Analysis and Factor Analysis;
3. Introduction to machine learning methods.

(All user inputs for this part should be selected in the code chunk in the raw .Rmd file) 

```{r setupfactor, echo=TRUE, tidy=TRUE}
# Please ENTER the original raw attributes to use. 
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
factor_attributes_used = c(2:30)

# Please ENTER the selection criteria for the factors to use. 
# Choices: "eigenvalue", "variance", "manual"
factor_selectionciterion = "manual"

# Please ENTER the desired minumum variance explained 
# (Only used in case "variance" is the factor selection criterion used). 
minimum_variance_explained = 65  # between 1 and 100

# Please ENTER the number of factors to use 
# (Only used in case "manual" is the factor selection criterion used).
manual_numb_factors_used = 6

# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quatimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Default is "varimax"
rotation_used = "varimax"

```

```{r}
factor_attributes_used <- intersect(factor_attributes_used, 1:ncol(ProjectData))
ProjectDataFactor <- ProjectData[,factor_attributes_used]
ProjectDataFactor <- data.matrix(ProjectDataFactor)
```

## Steps 1-2: Check the Data 

Start by some basic visual exploration of, say, a few data:

```{r}
rownames(ProjectDataFactor) <- paste0("Obs.", sprintf("%02i", 1:nrow(ProjectDataFactor)))
iprint.df(t(head(round(ProjectDataFactor, 2), max_data_report)))
```

The data we use here have the following descriptive statistics: 

```{r}
iprint.df(round(my_summary(ProjectDataFactor), 2))
```

## Step 3: Check Correlations

This is the correlation matrix of the customer responses to the `r ncol(ProjectDataFactor)` attitude questions - which are the only questions that we will use for the segmentation (see the case):

```{r}
thecor = round(cor(ProjectDataFactor),2)
iprint.df(round(thecor,2), scale=TRUE)
write.csv(round(thecor,2), file = "thecor.csv")
```

**Questions**

1. Do you see any high correlations between the responses? Do they make sense? 
2. What do these correlations imply?

**Answers:**

1
*When we see a high correlation between two of the questions, we understand that when a participant answers with a high value on one question, it is likely it will also answer with a high value on the other. This allows us to build a profile of the participants.  
*Looking at the correlations between the different questions, these seem to make sense. The ones that have high correlations with each other seem to potentially fit the same type of customer profile.

2
*Specifically, we have identified 3 groups of customers, by observing high correlations between three groups of questions. We have named the 3 groups of participants as: boat lovers, status seeking boat owners, leisure boat owners. We will characterize each of the three groups and identify the group of correlated questions for each.
*Boat lovers: these are the boat owners that are truly passionate about the sport, and have a deeper technical understanding of the boat. We can identify them most clearly through questions 15, 16, 17, 20, 23, 27, and 28. We see that these questions are positively correlated between themselves. They indicate that this type of customer is serious about the technology of the boat, that they are highly knowledgeable, and that they are truly passionate about boating.
*Status seeking boat owners: these are the type of customers that would buy a boat because of the status signal it provides, the prestige associated with owning a boat. We can identify them most clearly through questions 3, 5, 6, 7, 9, 13, and 14. They indicate that this type of client really values that the boat brand is recognized in the market, that they want to buy the latest and greatest, that they see their boat as a status symbol, and that owning a boat provides a sense of achievement.
*Leisure boat owners: these are the type of customers that own a boat in order to have a nice leisure activity in their free time. We can identify them most clearly through questions 18, 19, 21, 22, 24, 25, and 29. They indicate that this type of customer wants a boat in order to experience adventure and nature, to stay active and relax, and to socialize (family, friends).


## Step 4: Choose number of factors

Clearly the survey asked many redundant questions (can you think some reasons why?), so we may be able to actually "group" these 29 attitude questions into only a few "key factors". This not only will simplify the data, but will also greatly facilitate our understanding of the customers.

To do so, we use methods called [Principal Component Analysis](https://en.wikipedia.org/wiki/Principal_component_analysis) and [factor analysis](https://en.wikipedia.org/wiki/Factor_analysis) as also discussed in the [Dimensionality Reduction readings](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html). We can use two different R commands for this (they make slightly different information easily available as output): the command `principal` (check `help(principal)` from R package [psych](http://personality-project.org/r/psych/)), and the command `PCA` from R package [FactoMineR](http://factominer.free.fr) - there are more packages and commands for this, as these methods are very widely used.  

```{r}
# Here is how the `principal` function is used 
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Comp",1:ncol(UnRotated_Factors),sep="")
```

```{r}
# Here is how we use the `PCA` function 
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table

rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table), sep=" ")
colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")
```

Let's look at the **variance explained** as well as the **eigenvalues** (see session readings):

```{r}
iprint.df(round(Variance_Explained_Table, 2))
write.csv(round(Variance_Explained_Table,2), file = "Variance_Explained_Table.csv")

```

```{r}
eigenvalues  <- Variance_Explained_Table[, "Eigenvalue"]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
iplot.df(melt(df, id="components"))
```

**Questions:**

1. Can you explain what this table and the plot are? What do they indicate? What can we learn from these?
2. Why does the plot have this specific shape? Could the plotted line be increasing? 
3. What characteristics of these results would we prefer to see? Why?

**Answers**

1
*The table above shows us the eigen values and the percentage of variance explained for each of the components identified through the analysis. The higher the eigen value, the better it is at capturing the variance. There is a direct relationship between the eigen values and the percentage of variance explained: if one divides the eigen value by the number of components, one arrives at the percentage of variance explained.
*The plot shows us the eigen values plotted. It always goes from the component with the highest eigen value to the lowest. It allows for a simple visualization of how the eigen values gradually go down, and of the "marginal benefit" of adding an additional component. 
*We want to explain as much of the variance as possible, with as few factors as possible.
*There are three criteria we can use to choose the number of factors to be used. We can select all factors with an eigen value above 1, as these will have a strength above average. We can select the factors with highest eigen values, and define a threshold for the amount of variance we want to capture (e.g., 65%). Alternatively, we can select factors up to the "elbow" of the eigen values plot.

2
*As mentioned above, the plot has this shape because the components are organized in descending order, from the one with the highest eigen value to the one with the lowest (from the one that captures the most variance to the one that captures the least). This allows for a simple visualization of the "marginal benefit" of adding an additional component.
*It would not be possible for the plot to be increasing for the reason mentioned above.

3
*I think the results do not provide a very clear solution to the problem of how many factors one wants to choose. If we follow the first criteria (eigen value > 1), we would choose 5 factors. If we use the second criteria (cummulative percentage variance captured > 65%), we would choose to have 10 factors. Finally, the last criteria of the "elbow" of the eigen values plot is hard to use, since there is no clear "elbow". Using this last criteria we could perhaps use 3 factors, or 6 factors.

## Step 5: Interpret the factors

Let's now see how the "top factors" look like. 

```{r}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used
```

To better visualize them, we will use what is called a "rotation". There are many rotation methods. In this case we selected the `r rotation_used` rotation. For our data, the `r factors_selected` selected factors look as follows after this rotation: 

```{r}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Comp.",1:ncol(Rotated_Factors),sep="")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

iprint.df(Rotated_Factors, scale=TRUE)
write.csv(Rotated_Factors, file = "Rotated_Factors.csv")

```

To better visualize and interpret the factors we often "suppress" loadings with small values, e.g. with absolute values smaller than 0.5. In this case our factors look as follows after suppressing the small numbers:

```{r}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

iprint.df(Rotated_Factors_thres, scale=TRUE)
write.csv(Rotated_Factors_thres, file = "Rotated_Factors_thres.csv")

```

**Questions**

1. What do the first couple of factors mean? Do they make business sense? 
2. How many factors should we choose for this data/customer base? Please try a few and explain your final choice based on a) statistical arguments, b) on interpretation arguments, c) on business arguments (**you need to consider all three types of arguments**)
3. How would you interpret the factors you selected?
4. What lessons about data science do you learn when doing this analysis? Please comment. 

**Answers**

1
*The first table computed shows five factors (criteria eigenvalue > 1). They seem to make business sense. We have classified each of the factors in the order of the table:
  1.Purpose for ownership:  the questions that are associated with this factor, in order of importance, are 9, 26, 3, 13, 27, 28, 14, 16, 7, 17, and 20. Looking at these questions, we realize that they are mostly related to whether the customer values the status and prestige associated with owning a boat, or if they are instead truly passionate about the sport.
  2. Boat for leisure purposes: the questions that are associated with this factor, in order of importance, are 18, 21, 24, 19, 6, 22, and 7. Looking at these questions, we realize that they are mostly related to different usages of the boat (e.g., boat as an adventurous activity, as a way to experience nature, as a vehicle for socialization, ...).
  3. Value placed on quality: the questions that are associated with this factor, in order of importance, are 4, 5, and 15. All of these are related to the brand reputation and the quality of the boat/ technology.
  4. Customer knowledge on boats: the questions that are associated with this factor, in order of importance, are 11, 16, and 17. These all reflect the level of knowledge the customer has about boats - either because one is more knowledgeable than peers, or indicated by the ability to perform minor repairs and maintenance on own boat.
  5. Pricing: the questions that are associated with this factor, in order of importance, are 2, 12, and 8. These questions seem to reflect how the customer perceives the relationship between pricing and brand/ quality/ style.

2 + 3
*In order to understand the number of factors we should include in our analysis, we will test three different scenarios: choose factors with eigen value > 1, with cummulative variance of 65%+ explained, and according to the "elbow" of the eigen values plot.
*Eigen value >1
  - Number of factors: 5
  a) Statistical arguments
  We are using a statistical threshold to decide on 5 factors, as these are the ones that have a strength above average.
  b) Interpretation arguments
  We know that, in order for a factor to look "good", there are three quality criteria: 
    1. For each factor only a few loadings are large
    2. For each raw attribute only a few loadings are large (ideally only one)
    3. Any pair of factors should have different "patterns" of loading
  Here, we see that factor one has many loadings that are very large (11 - about 1/3 of the sample), so the first criteria is not met. For most of the raw attributes, only 1 or max 2 loadings are large. And the factors have different patterns among themselves. 
  c) Business arguments
  Looking at the factors, these all make sense from a business perspective, as mentioned above.
*Cumulative variance explained > 65%
  - Number of factors: 10
  a) Statistical arguments
  We are using a statistical threshold of 65% cumulative variance explained (a reasonable statistical threshold).
  b) Interpretation arguments
  From our three quality criteria, the first two criteria are well satisfied. However, regarding the patterns of loading, from factor 6 onwards it seems that we start to see a pattern that the factors have only 1 to 2 questions associated with that factor in a relevant way. This may be an indication that we are starting to have too many factors.
  c) Business arguments
  Factor 1: (Q27, 28, 20, 16, 23, 17) - Level of passion for boating
  Factor 2: (Q18, 21, 19, 24, 22) - Boat for leisure purposes
  Factor 3: (Q14, 13, 15) - Value attributed to latest model/ technology
  Factor 4: (Q7, 9, 3, 6) - Value attributed to status/ prestige
  Factor 5: (Q4, 2, 5) - Value attributed to brand reputation
  Factor 6: (Q11) - Repairing of own boat - this does not seem as a factor as relevant as the previous ones from a business perspective
  Factor 7: (Q10, 12) - these questions look at both knowledge of boats and preference for more/ less basic boat - this factor does not seem to make a lot of sense from a business perspective
  Factor 8: (Q8) - Value attributed to functionality vs style
  Factor 9: (Q1) - Level of shops visited
  Factor 10: (Q29) - this factor seems less relevant from a business perspective (too specific)
*Eigen values plot "elbow"
  - Number of factors: 6
  a) Statistical arguments
  Usage of "elbow" in eigen values plot as statistical threshold
  b) Interpretation arguments
  All three quality criteria seem to be fulfilled.
  c) Business arguments
  Factor 1: (Q21, 18, 19, 24, 6, 22) - Boat for leisure purposes (explained above)
  Factor 2: (Q27, 28, 16, 17, 20, 23) - Level of passion for boating
  Factor 3: (Q9, 3, 26, 7, 13, 14) - Level of value attributed to status/ prestige
  Factor 4: (Q4, 5) - Value attributed to brand reputation
  Factor 5: (Q8, 10) - Functionality vs style, and level of knowledge - difficult to interpret from business perspective
  Factor 6: (Q11) - Repairing of own boat - this does not seem as a factor as relevant as the previous ones from a business perspective 
*Considering the three models, model 3 seems to be the one that fulfills the three relevant criteria in a more satisfying way, since the first model aggregates too much on one factor, and model 2 starts to have too many non-relevant factors. 

4
*We learned that data science is a process that requires a lot of iteration and a combination of different criteria, in order to achieve robust results. In this case it is not possible to choose the most adequate number of factors, without trying different options, and qualifying them against different types of criteria.

## Step 6:  Save factor scores 

We can now either replace all initial variables used in this part with the factor scores, or just select one of the initial variables for each of the selected factors in order to represent that factor. Here is how the factor scores  are for the first few respondents:

```{r}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("DV (Factor)",1:ncol(NEW_ProjectData),sep=" ")

iprint.df(t(head(NEW_ProjectData, 10)), scale=TRUE)
write.csv(NEW_ProjectData, file = "FactorScores.csv")

```

**Questions**

1. Can you describe some of the people using the new derived variables (factor scores)? 
2. Which of the 29 initial variables would you select to represent each of the factors you selected?  

**Answers**

1
*Yes, the factors can help us understand the type of customer we are facing.
*For example, we can see that customer 1 cares a lot about brand reputation and status (factors 4 and 3 are the highest).
*Customer 7 on the other hand, seems to be more of a boat lover, and also has a higher than average ability to make fixes on his own boat (which makes sense since he will probably be more knowledgeable).

2
*For each of the factors I would select the following questions:
- Factor 1: Q18, 19, 21, 22, 24, 29
- Factor 2: Q11, 15, 16, 17, 20, 23, 27, 28
- Factor 3: Q3, 4, 5, 6, 7, 9, 13, 14
- Factor 4: Q3, 4, 5
- Factor 5: Q8, 10
- Factor 6: Q11

<hr>\clearpage

# Part 2: Customer Segmentation 

The code used here is along the lines of the code in the reading  [ClusterAnalysisReading.Rmd](https://github.com/InseadDataAnalytics/INSEADAnalytics/blob/master/CourseSessions/Sessions45/ClusterAnalysisReading.Rmd). We follow the process described in the [Cluster Analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions45/ClusterAnalysisReading.html) reading. 

In this part we also become familiar with:

1. Some clustering Methods;
2. How these tools can be used in practice.

A key family of methods used for segmentation is what is called **clustering methods**. This is a very important problem in statistics and **machine learning**, used in all sorts of applications such as in [Amazon's pioneer work on recommender systems](http://www.cs.umd.edu/~samir/498/Amazon-Recommendations.pdf). There are many *mathematical methods* for clustering. We will use two very standard methods, **hierarchical clustering** and **k-means**. While the "math" behind all these methods can be complex, the R functions used are relatively simple to use, as we will see. 

(All user inputs for this part should be selected in the code chunk in the raw .Rmd file) 

```{r setupcluster, echo=TRUE, tidy=TRUE}
# Please ENTER then original raw attributes to use for the segmentation (the "segmentation attributes")
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
segmentation_attributes_used = c(10,19,5,23) #c(10,19,5,12,3) 

# Please ENTER then original raw attributes to use for the profiling of the segments (the "profiling attributes")
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
profile_attributes_used = c(2:82) 

# Please ENTER the number of clusters to eventually use for this report
numb_clusters_used = 3 # for boats possibly use 5, for Mall_Visits use 3

# Please enter the method to use for the segmentation:
profile_with = "hclust" #  "hclust" or "kmeans"

# Please ENTER the distance metric eventually used for the clustering in case of hierarchical clustering 
# (e.g. "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski" - see help(dist)). 
# DEFAULT is "euclidean"
distance_used = "euclidean"

# Please ENTER the hierarchical clustering method to use (options are:
# "ward.D", "ward.D2", "single", "complete", "average", "mcquitty", "median" or "centroid").
# DEFAULT is "ward"
hclust_method = "ward.D"

# Please ENTER the kmeans clustering method to use (options are:
# "Hartigan-Wong", "Lloyd", "Forgy", "MacQueen").
# DEFAULT is "Lloyd"
kmeans_method = "Lloyd"

```

```{r}
# Same as the initial data
ProjectData <- ProjectData_INITIAL

segmentation_attributes_used <- intersect(segmentation_attributes_used, 1:ncol(ProjectData))
profile_attributes_used <- intersect(profile_attributes_used, 1:ncol(ProjectData))

ProjectData_segment <- ProjectData[,segmentation_attributes_used]
ProjectData_profile <- ProjectData[,profile_attributes_used]

ProjectData_scaled <- apply(ProjectData, 2, function(r) if (sd(r)!=0) (r-mean(r))/sd(r) else 0*r)
```

## Steps 1-2: Explore the data

(This was done above, so we skip it)

## Step 3. Select Segmentation Variables

For simplicity will use one representative question for each of the factor we found in Part 1 (we can also use the "factor scores" for each respondent) to represent our survey respondents. These are the `segmentation_attributes_used` selected below. We can choose the question with the highest absolute factor loading for each factor. For example, when we use 5 factors with the varimax rotation we can select questions Q.1.9 (I see my boat as a status symbol), Q1.18 (Boating gives me a feeling of adventure), Q1.4 (I only consider buying a boat from a reputable brand), Q1.11 (I tend to perform minor boat repairs and maintenance on my own) and Q1.2 (When buying a boat  getting the lowest price is more important than the boat brand) - try it. These are columns 10, 19, 5, 12, and 3, respectively of the data matrix `Projectdata`. 

## Step 4: Define similarity measure

We need to define a distance metric that measures how different people (observations in general) are from each other. This can be an important choice. Here are the differences between the observations using the distance metric we selected:

```{r}
euclidean_pairwise <- as.matrix(dist(head(ProjectData_segment, max_data_report), method="euclidean"))
euclidean_pairwise <- euclidean_pairwise*lower.tri(euclidean_pairwise) + euclidean_pairwise*diag(euclidean_pairwise) + 10e10*upper.tri(euclidean_pairwise)
euclidean_pairwise[euclidean_pairwise==10e10] <- NA
rownames(euclidean_pairwise) <- colnames(euclidean_pairwise) <- sprintf("Obs.%02d", 1:max_data_report)

iprint.df(round(euclidean_pairwise))
```

## Step 5: Visualize Pair-wise Distances

We can see the histogram of, say, the first 2 variables (can you change the code chunk in the raw .Rmd file to see other variables?)

```{r}
variables_to_plot = 1:3
do.call(iplot.grid, lapply(variables_to_plot, function(n){
  iplot.hist(ProjectData_segment[, n], breaks=10, xlab = paste("Variable", n))
}))
```

or the histogram of all pairwise distances for the `r distance_used` distance:

```{r}
Pairwise_Distances <- dist(ProjectData_segment, method = distance_used) 
iplot.hist(Pairwise_Distances, breaks=10)
```

## Step 6: Method and Number of Segments

We need to select the clustering method to use, as well as the number of cluster. It may be useful to see the dendrogram from Hierarchical Clustering, to have a quick idea of how the data may be segmented and how many segments there may be. Here is the dendrogram for our data:

```{r}
Hierarchical_Cluster_distances <- dist(ProjectData_segment, method=distance_used)
Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=hclust_method)
# Display dendogram
iplot.dendrogram(Hierarchical_Cluster)
# TODO: Draw dendogram with red borders around the 3 clusters
#rect.hclust(Hierarchical_Cluster, k=5, border="red") 
```

We can also plot the "distances" traveled before we need to merge any of the lower and smaller in size clusters into larger ones - the heights of the tree branches that link the clusters as we traverse the tree from its leaves to its root. If we have n observations, this plot has n-1 numbers, we see the first 20 here. 
```{r}
num <- nrow(ProjectData) - 1
df1 <- cbind(as.data.frame(Hierarchical_Cluster$height[length(Hierarchical_Cluster$height):1]), c(1:num))
colnames(df1) <- c("distances","index")
iplot.df(melt(head(df1, 20), id="index"), xlab="Number of Components")
```

Here is the segment membership of the first `r max_data_report` respondents if we use hierarchical clustering:

```{r}
cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=5)) # cut tree into as many clusters as numb_clusters_used
cluster_ids_hclust=unique(cluster_memberships_hclust)

ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_hclust_membership, max_data_report), 2))
write.csv(round(ProjectData_with_hclust_membership, 2), file = "ProjectData_with_hclust_membership.csv")

```

while this is the segment membership if we use k-means:

```{r}
kmeans_clusters <- kmeans(ProjectData_segment,centers= numb_clusters_used, iter.max=2000, algorithm=kmeans_method)

ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_kmeans_membership, max_data_report), 2))
write.csv(round(ProjectData_with_kmeans_membership, 2), file = "ProjectData_with_kmeans_membership.csv")

```

## Step 7: Profile and interpret the segments 

In market segmentation one may use variables to **profile** the segments which are not the same (necessarily) as those used to **segment** the market: the latter may be, for example, attitude/needs related (you define segments based on what the customers "need"), while the former may be any information that allows a company to identify the defined customer segments (e.g. demographics, location, etc). Of course deciding which variables to use for segmentation and which to use for profiling (and then **activation** of the segmentation for business purposes) is largely subjective.  In this case we can use all survey questions for profiling for now - the `profile_attributes_used` variables selected below. 

There are many ways to do the profiling of the segments. For example, here we show how the *average* answers of the respondents *in each segment* compare to the *average answer of all respondents* using the ratio of the two.  The idea is that if in a segment the average response to a question is very different (e.g. away from ratio of 1) than the overall average, then that question may indicate something about the segment relative to the total population. 

Here are for example the profiles of the segments using the clusters found above.  First let's see just the average answer people gave to each question for the different segments as well as the total population:

```{r}
cluster_memberships_kmeans <- kmeans_clusters$cluster 
cluster_ids_kmeans <- unique(cluster_memberships_kmeans)

if (profile_with == "hclust"){
  cluster_memberships <- cluster_memberships_hclust
  cluster_ids <-  cluster_ids_hclust  
}
if (profile_with == "kmeans"){
  cluster_memberships <- cluster_memberships_kmeans
  cluster_ids <-  cluster_ids_kmeans
}

# WE WILL USE THESE IN THE CLASSIFICATION PART LATER
NewData = matrix(cluster_memberships,ncol=1)

population_average = matrix(apply(ProjectData_profile, 2, mean), ncol=1)
colnames(population_average) <- "Population"
Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profile[(cluster_memberships==i), ], 2, mean))
if (ncol(ProjectData_profile) <2)
  Cluster_Profile_mean=t(Cluster_Profile_mean)
colnames(Cluster_Profile_mean) <- paste("Seg.", 1:length(cluster_ids), sep="")
cluster.profile <- cbind (population_average,Cluster_Profile_mean)

iprint.df(round(cluster.profile, 2))
write.csv(round(cluster.profile, 2), file = "cluster.profile.csv")

```

We can also "visualize" the segments using **snake plots** for each cluster. For example, we can plot the means of the profiling variables for each of our clusters to better visualize differences between segments. For better visualization we plot the standardized profiling variables.

```{r}
ProjectData_scaled_profile = ProjectData_scaled[, profile_attributes_used,drop=F]

Cluster_Profile_standar_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_scaled_profile[(cluster_memberships==i), ,drop = F], 2, mean))
if (ncol(ProjectData_scaled_profile) < 2)
  Cluster_Profile_standar_mean = t(Cluster_Profile_standar_mean)
colnames(Cluster_Profile_standar_mean) <- paste("Seg ", 1:length(cluster_ids), sep="")

iplot.df(melt(cbind.data.frame(idx=as.numeric(1:nrow(Cluster_Profile_standar_mean)), Cluster_Profile_standar_mean), id="idx"), xlab="Profiling variables (standardized)",  ylab="Mean of cluster")

write.csv(round(Cluster_Profile_standar_mean, 2), file = "Cluster_Profile_standar_mean.csv")

```

We can also compare the averages of the profiling variables of each segment relative to the average of the variables across the whole population. This can also help us better understand whether  there are indeed clusters in our data (e.g. if all segments are much like the overall population, there may be no segments). For example, we can measure the ratios of the average for each cluster to the average of the population, minus 1, (e.g. `avg(cluster)` `/` `avg(population)` `-1`) for each segment and variable:

```{r}
population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
cluster_profile_ratios <- (ifelse(population_average_matrix==0, 0,Cluster_Profile_mean/population_average_matrix))
colnames(cluster_profile_ratios) <- paste("Seg.", 1:ncol(cluster_profile_ratios), sep="")
rownames(cluster_profile_ratios) <- colnames(ProjectData)[profile_attributes_used]
## printing the result in a clean-slate table
iprint.df(round(cluster_profile_ratios-1, 2))

```

**Questions**

1. What do the numbers in the last table indicate? What numbers are the more informative?
2. Based on the tables and snake plot above, what are some key features of each of the segments of this solution?

**Answers**

1. The numbers in the last table represent how similar or different the answers within that particular segment are from the average of all answers. This can then show us what differentiates each segment from the rest of the data. The nearer 0 the number is,the closer to the average answer that segment's responses are. However the higher the number, the further away that segments answers are to the average answers and therefore the more significant that question is in defining that particular segment.

2.
Segment 1:
Care about the brand and believe that the brand of boat says a lot about themselves
Sees the boat as a status symbol - this is the strongest stand out characteristic of this segment
Believe that boats help them to stay active

Segment 2:
Only considers buying boats from reputable brands
Do not see boats as a status symbol at all
Tends to buy the latest and greatest accessories for their boats
Boating gives a feeling of adventure
Being in charge of their boat is very important to them

Segment 3:
Having a low price is highly important
Only considers buying boats from reputable brands
Tends to prefer boats with few (little or no) frills
Boating gives a feeling of adventure

Segment 4:
Does not see the boat as a status symbol
Does not feel that the boat keeps them active
Does not feel that the boat gives a feeling of adventure

Segment 5:
The brand of boat says a lot about them
Owning a boat is a way of rewarding themseleves for hard work
Owning a boat gives a sense of achievement
ends to buy the latest and greatest accessories for their boats
Considers themselves more knowledgeable about boats than their friends
Boating helps them escape everyday lives


## Step 8: Robustness Analysis

We should also consider the robustness of our analysis as we change the clustering method and parameters. Once we are comfortable with the solution we can finally answer our first business questions: 

**Questions**

1. How many segments are there in our market? How many do you select and why? Try a few and explain your final choice based on a) statistical arguments, b) on interpretation arguments, c) on business arguments (**you need to consider all three types of arguments**)
2. Can you describe the segments you found based on the profiles?
3. What if you change the number of factors and in general you *iterate the whole analysis*? **Iterations** are key in data science.
4. Can you now answer the [Boats case questions](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-A-prerelease.pdf)? What business decisions do you recommend to this company based on your analysis?

**Answers**

1.There are 3 segments in our market because when categorised into 3 segments the groups are different enough from each other to be of use. This allows us to identify key attributes of those three groups and make business decisoins based on that. 

The three segments are separated into groups that:
1) Boats for status
2) Boats for adventure and most up to date technology
3) Boats for fun but cheap

These three segments cover 3 very different aspects of the market and have clear differentiating factors from a statistical perspective. This then enables the company to create different products to suit these customer segments adn to market them very differently, depending on the key attributes that those customers look for when boat shopping - as idenfied in the previous questions.

2. the three segments we are focusing on are 1,2 and 3:
1) Boats for status: these are people who want top brands to show their status and wealth, they are the top earners, generally in paid work and are willing to spend up to 50k on a boat
2) Boats for adventure and most up to date technology: these people want a good boat with latest technology but are focused on the adventure and sailing aspect. They are willing to spend a middling amount on the boat and are not interested in the status that the brand of the boat lends them.
3) Boats for fun but cheap: These people don't want to spend a huge amount but want to enjoy their time out on the boat in the cheapest, most fun and easy way possible

3. If you changed the number of factors you may be able to get more or less granular descriptions of the segments. By increasing the factors, this may enable you to split into more different sectors and therefore target each sector more specifically. Equally if you wanted to reduce the number of factors you may end up with fewer, wider defined segments and will be unable to specific your marketing and products to particualr groups as well.

4. Our recommnedation is to create 3 different products, targeting the 3 different customer segments:

Product 1: Flashy and high end, showing wealth and status and priced at the high end of the market.
Product 2: A speedy boat that is targeted at the true sailors who love the latest technology
Product 3: A paired back, simple boat with the bare minimum for hobbiests to enjoy cheaply. Priced at the lower end of the market

We would recommend marketing these boats very differently so as to target them at the specific customer segments.


<hr>\clearpage

# Part 3: Purchase Drivers

We will now use the [classification analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions67/ClassificationAnalysisReading.html) methods to understand the key purchase drivers for boats (a similar analysis can be done for recommendation drivers). For simplicity we do not follow the "generic" steps of classification discussed in that reading, and only consider the classification and purchase drivers analysis for the segments we found above. 

We are interested in understanding the purchase drivers, hence our **dependent** variable is column 82 of the Boats data (`r colnames(ProjectData)[82]`) - why is that? We will use only the subquestions of **Question 16** of the case for now, and also select some of the parameters for this part of the analysis: 

```{r setupclassification, echo=TRUE, tidy=TRUE}
# Please ENTER the class (dependent) variable:
# Please use numbers, not column names! e.g. 82 uses the 82nd column are dependent variable.
# YOU NEED TO MAKE SURE THAT THE DEPENDENT VARIABLES TAKES ONLY 2 VALUES: 0 and 1!!!
dependent_variable= 82

# Please ENTER the attributes to use as independent variables 
# Please use numbers, not column names! e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
independent_variables= c(54:80) # use 54-80 for boats

# Please ENTER the profit/cost values for the correctly and wrong classified data:
actual_1_predict_1 = 100
actual_1_predict_0 = -75
actual_0_predict_1 = -50
actual_0_predict_0 = 0

# Please ENTER the probability threshold above which an observations  
# is predicted as class 1:
Probability_Threshold=50 # between 1 and 99%

# Please ENTER the percentage of data used for estimation
estimation_data_percent = 80
validation_data_percent = 10

# Please enter 0 if you want to "randomly" split the data in estimation and validation/test
random_sampling = 0

# Tree parameter
# PLEASE ENTER THE Tree (CART) complexity control cp (e.g. 0.001 to 0.02, depending on the data)
CART_cp = 0.01

# Please enter the minimum size of a segment for the analysis to be done only for that segment
min_segment = 100
```

```{r}
ProjectData = ProjectData_INITIAL # Just to initialize the data

Probability_Threshold = Probability_Threshold/100 # make it between 0 and 1

dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(ProjectData), max(i,1))))
independent_variables = unique(sapply(independent_variables,function(i) min(ncol(ProjectData), max(i,1))))

if (length(unique(ProjectData[,dependent_variable])) !=2){
  cat("\n*****\n BE CAREFUL, THE DEPENDENT VARIABLE TAKES MORE THAN 2 VALUES...")
  cat("\nSplitting it around its median...\n*****\n ")
  new_dependent = ProjectData[,dependent_variable] >= median(ProjectData[,dependent_variable])
  ProjectData[,dependent_variable] <- 1*new_dependent
}

Profit_Matrix = matrix(c(actual_1_predict_1, actual_0_predict_1, actual_1_predict_0, actual_0_predict_0), ncol=2)
colnames(Profit_Matrix)<- c("Predict 1", "Predict 0")
rownames(Profit_Matrix) <- c("Actual 1", "Actual 0")
test_data_percent = 100-estimation_data_percent-validation_data_percent
CART_control = rpart.control(cp = CART_cp)
```

**Questions**

1. How do you select the profit/cost values for the analysis? Does the variable `r Profit_Matrix` above relate to the final business decisions? How?
2. What does the variable "Probability_Threshold" affect? Does it relate to the final business decisions? How?

**Answers**

*
*
*
*
*
*
*
*
*
*

We will use two classification trees and logistic regression. You can select "complexity" control for one of the classification trees in the code chunk of the raw .Rmd file here

```{r CART_control, echo=TRUE, tidy=TRUE}
CART_control = 0.001
```

**Question**

1. How can this parameter affect the final results? What business implications can this parameter choice have?

**Answer**

*
*
*
*
*
*
*
*
*
*

This is a "small tree" classification for example:

```{r}
# FIrst we split the data in estimation, validation, and test

if (random_sampling){
  estimation_data_ids=sample.int(nrow(ProjectData),floor(estimation_data_percent*nrow(ProjectData)/100))
  non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
  validation_data_ids=non_estimation_data[sample.int(length(non_estimation_data), floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))]
  } else {
    estimation_data_ids=1:floor(estimation_data_percent*nrow(ProjectData)/100)
    non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
    validation_data_ids = (tail(estimation_data_ids,1)+1):(tail(estimation_data_ids,1) + floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))
    }

test_data_ids = setdiff(1:nrow(ProjectData), union(estimation_data_ids,validation_data_ids))

estimation_data=ProjectData[estimation_data_ids,]
validation_data=ProjectData[validation_data_ids,]
test_data=ProjectData[test_data_ids,]
```

```{r}
# just name the variables numerically so that they look ok on the tree plots
independent_variables_nolabel = paste("IV", 1:length(independent_variables), sep="")

estimation_data_nolabel = cbind(estimation_data[,dependent_variable], estimation_data[,independent_variables])
colnames(estimation_data_nolabel)<- c(colnames(estimation_data)[dependent_variable],independent_variables_nolabel)

validation_data_nolabel = cbind(validation_data[,dependent_variable], validation_data[,independent_variables])
colnames(validation_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

test_data_nolabel = cbind(test_data[,dependent_variable], test_data[,independent_variables])
colnames(test_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

estimation_data_nolabel = data.frame(estimation_data_nolabel)
validation_data_nolabel = data.frame(validation_data_nolabel)
test_data_nolabel = data.frame(test_data_nolabel)

estimation_data = data.frame(estimation_data)
validation_data = data.frame(validation_data)
test_data = data.frame(test_data)
```

```{r}
formula=paste(colnames(estimation_data)[dependent_variable],paste(Reduce(paste,sapply(head(independent_variables_nolabel,-1), function(i) paste(i,"+",sep=""))),tail(independent_variables_nolabel,1),sep=""),sep="~")
CART_tree<-rpart(formula, data= estimation_data_nolabel,method="class", control=CART_control)

rpart.plot(CART_tree, box.palette="OrBu", type=3, extra=1, fallen.leaves=F, branch.lty=3)
```

```{r}
CART_tree_large<-rpart(formula, data= estimation_data_nolabel,method="class", control=rpart.control(cp = 0.005))
```

```{r}
# Let's first calculate all probabilites for the estimation, validation, and test data
estimation_Probability_class1_tree<-predict(CART_tree, estimation_data_nolabel)[,2]
estimation_Probability_class1_tree_large<-predict(CART_tree_large, estimation_data_nolabel)[,2]

validation_Probability_class1_tree<-predict(CART_tree, validation_data_nolabel)[,2]
validation_Probability_class1_tree_large<-predict(CART_tree_large, validation_data_nolabel)[,2]

test_Probability_class1_tree<-predict(CART_tree, test_data_nolabel)[,2]
test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_nolabel)[,2]


estimation_prediction_class_tree=1*as.vector(estimation_Probability_class1_tree > Probability_Threshold)
estimation_prediction_class_tree_large=1*as.vector(estimation_Probability_class1_tree_large > Probability_Threshold)

validation_prediction_class_tree=1*as.vector(validation_Probability_class1_tree > Probability_Threshold)
validation_prediction_class_tree_large=1*as.vector(validation_Probability_class1_tree_large > Probability_Threshold)

test_prediction_class_tree=1*as.vector(test_Probability_class1_tree > Probability_Threshold)
test_prediction_class_tree_large=1*as.vector(test_Probability_class1_tree_large > Probability_Threshold)
```

```{r}
formula_log=paste(colnames(estimation_data[,dependent_variable,drop=F]),paste(Reduce(paste,sapply(head(independent_variables,-1), function(i) paste(colnames(estimation_data)[i],"+",sep=""))),colnames(estimation_data)[tail(independent_variables,1)],sep=""),sep="~")

logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data)

log_coefficients = round(summary(logreg_solution)$coefficients,1)
```

```{r}
estimation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=estimation_data[,independent_variables])
validation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=validation_data[,independent_variables])
test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data[,independent_variables])

estimation_prediction_class_log=1*as.vector(estimation_Probability_class1_log > Probability_Threshold)
validation_prediction_class_log=1*as.vector(validation_Probability_class1_log > Probability_Threshold)
test_prediction_class_log=1*as.vector(test_Probability_class1_log > Probability_Threshold)
```

After also running the large tree and the logistic regression classifiers, we can then check how much "weight" these three methods put on the different purchase drivers (Q16 of the survey):

```{r}
log_importance = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
log_importance = log_importance/max(abs(log_importance))

tree_importance = CART_tree$variable.importance
tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
tree_importance_final = rep(0,length(independent_variables))
tree_importance_final[tree_ordered_drivers] <- tree_importance
tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
tree_importance_final <- tree_importance_final*sign(log_importance)

large_tree_importance = CART_tree_large$variable.importance
large_tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree_large$variable.importance)))
large_tree_importance_final = rep(0,length(independent_variables))
large_tree_importance_final[large_tree_ordered_drivers] <- large_tree_importance
large_tree_importance_final <- large_tree_importance_final/max(abs(large_tree_importance_final))
large_tree_importance_final <- large_tree_importance_final*sign(log_importance)

Importance_table <- cbind(tree_importance_final,large_tree_importance_final, log_importance)
colnames(Importance_table) <- c("CART 1", "CART 2", "Logistic Regr.")
rownames(Importance_table) <- rownames(log_importance)

iprint.df(Importance_table)
```

Finally, if we were to use the estimated classification models on the test data, we would get the following profit curves (see the raw .Rmd file to select the business profit parameters). 

The profit curve using the small classification tree: 

```{r}
actual_class<- test_data[,dependent_variable]

probs = test_Probability_class1_tree
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_small_tree = df[which.max(df$Profit),]
```

The profit curve using the large classification tree: 

```{r}
probs = test_Probability_class1_tree_large
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_large_tree = df[which.max(df$Profit),]
```

The profit curve using the logistic regression classifier: 

```{r}
probs = test_Probability_class1_log
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_logistic = df[which.max(df$Profit),]
```

These are the maximum total profit achieved in the test data using the three classifiers (without any segment specific analysis so far).

```{r}
best_profits = rbind(best_profits_small_tree, best_profits_large_tree, best_profits_logistic)
rownames(best_profits) <- c("Small Tree", "Large Tree", "Logistic Regression")
iprint.df(round(best_profits, 2))
```

<hr>\clearpage

# Part 4: Business Decisions

We will now get the results of the overall process (parts 1-3) and based on them make business decisions (e.g. answer the questions of the Boats case study). Specifically, we will study the purchase drivers for each segment we found and consider the profit curves of the developed models on our test data. 

**Final Solution: Segment Specific Analysis**

Let's see first how many observations we have in each segment, for the segments we selected above:

```{r}
# Let's rename cluster_memberships as cluster_ids
cluster_ids <- cluster_memberships

cluster_size = NULL
for (i in sort(unique(cluster_ids))){
  cluster_size = c(cluster_size,sum(cluster_ids == i))
  }
cluster_size = matrix(cluster_size, nrow=1)
colnames(cluster_size) <- paste("Segment", 1:length(cluster_size), sep=" ")
rownames(cluster_size) <- "Number of Obs."
iprint.df(cluster_size, scale=TRUE)
```

This is our final segment specific analysis and solution. We can study now the purchase drivers (average answers to Q16 of the survey) for each segment. They are as follows:

```{r}
actual_class<- test_data[,dependent_variable]
probs_tree = 0*test_Probability_class1_tree
probs_tree_large = 0*test_Probability_class1_tree_large
probs_log = 0*test_Probability_class1_log
Log_Drivers = NULL

for (i in sort(unique(cluster_ids))){
  useonly = which(cluster_ids==i)
  if (length(useonly) >= min_segment){
    
    test_ids_used = intersect(test_data_ids,useonly)
    probs_to_fill = which(sapply(test_data_ids, function(i) sum(test_ids_used==i)) !=0)
    estimation_data_clus=ProjectData[intersect(estimation_data_ids,useonly) ,]
    test_data_clus=ProjectData[intersect(test_data_ids,useonly),]
    
    ###
    estimation_data_clus_nolabel = cbind(estimation_data_clus[,dependent_variable], estimation_data_clus[,independent_variables])
    colnames(estimation_data_clus_nolabel)<- c(colnames(estimation_data_clus)[dependent_variable],independent_variables_nolabel)
    
    test_data_clus_nolabel = cbind(test_data_clus[,dependent_variable], test_data_clus[,independent_variables])
    colnames(test_data_clus_nolabel)<- c(dependent_variable,independent_variables_nolabel)
    
    estimation_data_clus = data.frame(estimation_data_clus)
    test_data_clus = data.frame(test_data_clus)
    estimation_data_clus_nolabel = data.frame(estimation_data_clus_nolabel)
    test_data_clus_nolabel = data.frame(test_data_clus_nolabel)
    ###
    
    CART_tree<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=CART_control)    
    CART_tree_large<-rpart(formula, data= estimation_data_clus_nolabel,method="class", control=rpart.control(cp = 0.005))
    logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data_clus)
    
    #####
    
    test_Probability_class1_tree<-predict(CART_tree, test_data_clus_nolabel)[,2]
    test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_clus_nolabel)[,2]
    test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data_clus[,independent_variables])
    
    #######
    probs_tree[probs_to_fill] <- test_Probability_class1_tree
    probs_tree_large[probs_to_fill] <- test_Probability_class1_tree
    probs_log[probs_to_fill] <- test_Probability_class1_log
    
    
    log_coefficients = round(summary(logreg_solution)$coefficients,1)
    Log_Drivers_segment = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
    Log_Drivers_segment = Log_Drivers_segment/max(abs(Log_Drivers_segment))
    
    tree_importance = CART_tree$variable.importance
    tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
    tree_importance_final = rep(0,length(independent_variables))
    tree_importance_final[tree_ordered_drivers] <- tree_importance
    tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
    tree_importance_final <- tree_importance_final*sign(Log_Drivers_segment)
    
    #Log_Drivers = cbind(Log_Drivers,tree_importance_final)
    Log_Drivers = cbind(Log_Drivers,Log_Drivers_segment)

    
    }
  }
colnames(Log_Drivers) <- paste("Segment", 1:length(unique(cluster_ids)), sep = " ")
iprint.df(round(tail(Log_Drivers,-1), 2))
```

The profit curves for the test data in this case are as follows. The profit curve using the small classification tree is: 

```{r}
actual_class<- test_data[,dependent_variable]

probs = probs_tree
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_small_tree = df[which.max(df$Profit),]
```

The profit curve using the large classification tree is: 

```{r}
probs = probs_tree_large
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_large_tree = df[which.max(df$Profit),]
```

The profit curve using the logistic regression classifier: 

```{r}
probs = probs_log
xaxis = sort(unique(c(0,1,probs)), decreasing = TRUE)
res = Reduce(cbind,lapply(xaxis, function(prob){
  useonly = which(probs >= prob)
  predict_class = 1*(probs >= prob)
  theprofit = Profit_Matrix[1,1]*sum(predict_class==1 & actual_class ==1)+
    Profit_Matrix[1,2]*sum(predict_class==0 & actual_class ==1)+
    Profit_Matrix[2,1]*sum(predict_class==1 & actual_class ==0)+
    Profit_Matrix[2,2]*sum(predict_class==0 & actual_class ==0)
  
  c(100*length(useonly)/length(actual_class), theprofit) 
  }))
xaxis = res[1,]; yaxis = res[2,]
names(xaxis)<- NULL; names(yaxis) <- NULL
df<-data.frame(Percentile = xaxis, Profit = yaxis)
iplot.df(df, x="Percentile", y="Profit", v=NULL)

best_profits_logistic = df[which.max(df$Profit),]
```

These are the maximum total profit achieved in the test data using the three classifiers with the selected market segmentation solution.

```{r}
best_profits = rbind(best_profits_small_tree, best_profits_large_tree, best_profits_logistic)
rownames(best_profits) <- c("Small Tree", "Large Tree", "Logistic Regression")
iprint.df(round(best_profits, 2))
```

**Questions:**

1. What are the main purchase drivers for the segments and solution you found?
2. How different are the purchase drivers you find when you use segmentation versus when you study all customers as "one segment"? Why? 
3. Based on the overall analysis, what segmentation would you choose?
4. What is the business profit the company can achieve (as measured with the test data) based on your solution? 
5. What business decisions can the company make based on this analysis?

**Answers:**

*
*
*
*
*
*
*
*
*
*

**You have now completed your first market segmentation project.** Do you have data from another survey you can use with this report now? 

**Extra question**: explore and report a new segmentation analysis... 